---
title: |
  | \vspace{5cm} \LARGE [Filler Title]
  | \vspace{0.5cm} \Large ECON 573 Group 6
author: "Jemin Park, Jonah Selom, Keegan McDowell"
date: "4/30/2022"
# setspace: doublespacing
fontsize: 12pt
documentclass: article
geometry: margin=3cm
indent: true
output:
  pdf_document:
    highlight: tango
    number_sections: true
header-includes:
  - \usepackage{setspace}\doublespacing
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}

---

```{r setup, include=FALSE}
# don't want to show code on default
knitr::opts_chunk$set(echo = FALSE) 
rm(list=ls())
```

\newpage
\maketitle
\newpage
\tableofcontents
\newpage


# Introduction 

Filler paragraph 1

Filler para 2

# Data

Discussion of data source

## Data Preprocessing

Discussion on cleaning and other stuff

```{r data-import}
# data import
1
```

## Data Visualization

(each into separate sections as needed)


# Methods

High level discussion on approach

## Model 1

Discussion & results

## Model 2

Discussion & results

## Model 3

Discussion & results

## Model 4

Discussion & results

## Model 5

Discussion & results

# Conclusion

Final discussion with interpretation from model, inferences, best results, etc.

